---
title: 'Predicting Bank Customer Behavior: A Data-Driven Approach to Campaign Effectiveness and Customer Segmentation'
author: 'Toyin'
format: 
    dashboard:
      logo: DULogo.png
      scrolling: true
      theme: united
embed-resources: true
---

```{r}
install.packages("xgboost")
```


```{r setup,  include=FALSE,warning=FALSE}
#include=FALSE will not include r code in output
#warning=FALSE will remove any warnings from output

library(ISLR) #v1.4 College data
library(flexdashboard) #v0.6.2
library(tidymodels) #v1.2.0
  #library(parsnip) #v1.2.1 linear_reg(), set_engine(), set_mode(), fit(), predict()
  #library(yardstick) #v1.3.1 metrics(), rac_auc(), roc_curve(), metric_set(), conf_matrix()
  #library(dplyr) #v1.1.4 %>%, select(), select_if(), filter(), mutate(), group_by(), 
    #summarize(), tibble()
  #library(ggplot2) #v3.5.1 ggplot()
  #library(broom) #v1.0.6 for tidy(), augment(), glance()
  #library(rsample) #v1.2.1 initial_split()
library(mgcv) #v1.9-1 gam() - different from book gam()
library(mgcViz) #v0.1.11 getViz() plot for gams
library(gridExtra) #v2.3 grid.arrange
library(knitr) #v1.48 kable(digits=) #formatted tables 
library(skimr)
library(janitor)
library(GGally)
library(ggplot2)
library(readr)
library(rmarkdown)
library(sjPlot)
library(vip)
library(performance)
library(recipes)
library(themis)
library(ranger)
library(grid)
library(ggcorrplot)
library(car)
library(GGally)
library(dplyr)
library(tidymodels)
library(glmnet)
library(tune)
library(workflows)
library(xgboost)
theme_set(theme_bw()) #sets default ggplot output style
```

# EXECUTIVE SUMMARY
## Row
::: {.card title="Executive Summary" fill="false"}
**Executive Summary**

The project aimed to build predictive models to understand the factors influencing individuals' account balances. Regression models, including Multiple Regression, Lasso Tuned Regression, and Generalized Additive Models (GAMs), were evaluated based on their ability to predict account balances.

After assessing the performance of the models, the Generalized Additive Model (GAMs) emerged as the best-performing model, with the lowest Root Mean Square Error (RMSE) and Mean Absolute Error (MAE), indicating it provided the most accurate predictions. While R-squared values for all models were low, suggesting limited explanatory power, GAMs still captured more variance than the other models even though the anount of variance it captured is super low. In comparison, both Multiple Regression and Lasso Tuned regression showed higher errors and slightly worse predictive accuracy.

Additionally, the Logistic Model, Tuned Random Forest and Tuned Boosted Models were also tested to predict the factors that influences individual's subscription to a product. Tuned Random Forest and Tuned Boosted Models showed excellent performance in distinguishing between subscribers and non-subscribers, achieving high accuracy, sensitivity, and specificity.

In conclusion, GAMs proved to be the most effective regression model for understanding subscription patterns, while Random Forest and Boosted Models excelled in classification tasks. Moving forward, the GAMs model can be used for predicting subscription likelihood but transformation needs to be done to the varibales as their distribution is not normal. In addition, some factors might need to be removed because they are probably making the model noisy while some key predictors might also need to be included. As for the classification models, further fine-tuning of the classification models may provide even better insights into the key factors driving subscriptions.

:::

# INTRODUCTION
## Row
### Column {width=60%}
::: {.card title="Introduction" fill="false"}


**The Problem Description**

These days, bank marketing has become a necessity for financial institutions to attract, secure, and maintain customers while promoting trust and financial confidence. It not only improves brand awareness and customer relationships but also ensures that banks remain competitive in a rapidly growing financial landscape. Through effective marketing strategies, banks can highlight their unique services, differentiate themselves from competitors, and foster long-term customer loyalty.
In this analysis, regression and classification models will be used to extract insights from a bank marketing dataset, identifying key factors influencing customer balances and predicting which customers are most likely to apply for a housing loan

**Project Goal:** 

The goal of this project is to utilize quantitative analysis (Multiple Linear Regression, Lasso Regression, and Generalized Additive Models) to identify and quantify which numerical campaign and customer-related factors (age, duration of campaign, frequency of contact) significantly influence customers’ average bank balances. Additionally, using qualitative (categorical) analysis methods—Multiple Logistic Regression, Tuned Random Forest, Tuned Boosted Trees, and Tuned Support Vector Machines—the project aims to determine which demographic and campaign-related qualitative factors (e.g., loan status, housing, marital status, and job type) significantly affect a customer's likelihood of subscribing to a term deposit.

The insights generated from these combined analyses will enable targeted marketing strategies, effective customer segmentation, and improved understanding of customer behavior for better financial and business decisions.

 
**Business Questions:**

1. Quantitative Analysis Questions (Target Variable: balance):

a. Which numerical demographic and campaign factors significantly predict the customer's average account balance?

b. How do quantitative factors (age of customer, duration of contact, frequency of campaigns, days since last contact, and previous interactions influence customers' bank balances?

c. Can we accurately predict a customer's bank balance using quantitative variables derived from demographic profiles and past campaign engagements?

 

2. Qualitative Analysis Questions (Target Variable: y - Subscription Decision):

a. What qualitative categorical factors most significantly influence a customer's decision to subscribe to a term deposit?

b. How do qualitative variables like loan status, housing status, marital status, and job type impact the likelihood of a customer subscribing to a term deposit?

c. Can we accurately classify customers as potential subscribers based on categorical demographic attributes and campaign-related factors?

:::

### Column {width=40%}
::: {.card title="Data"}
**DATA:**

* **- Variables: 16 total variables**
* **- Quantitative Variables: 6**
* **- Qualitative Variables: 10**
* **- Rows: 45,211**
 
**Quantitative Variables**

* **Age**: The individual's age	Job – The type of job the individual has
* **Balance**: The individual's bank balance	Marital – Marital status of the individual
* **Duration**: Duration of the last contact (in seconds)	Education – The individual's education level
* **Campaign**: Number of contacts performed during the current campaign	Default – Whether the individual defaulted on a loan (binary)
* **Previous**: Number of contacts performed before the current campaign	Loan – Whether the individual has a personal loan (binary)
* **Day**: Day of the last month the individual was contacted (numeric)	Contact – The communication type used to contact the individual

* **Qualitative Variables**

* **Job**: The type of job the individual has
* **Marital**: Marital status of the individual
* **Education**: The individual's education level
* **Default**: Whether the individual defaulted on a loan (binary)
* **Housing**: Whether the individual has a housing loan (binary)
* **Loan**: Whether the individual has a personal loan (binary)
* **Contact**: The communication type used to contact the individual
* **Month**: The month when the individual was last contacted
* **Poutcome**: Outcome of the previous marketing campaign
* **Y** – Whether the individual subscribed to the product (binary)

**We want to predict**

* **Balance**
* **y**

**DATA SOURCE:**
Moro, S., Cortez, P., Rita, P. (2014). A data-driven approach to predict the success of bank telemarketing. Decision Support Systems, Business Computer Economics. DOI:https://archive.ics.uci.edu/dataset/222/bank+marketing 

:::

Loading the dataset
```{r load_data}
# Loading the dataset and cleaning column names
fin_data <- read_csv("FinData.csv") %>%
  clean_names() %>%
  na.omit()

# For regression analysis 
regression_df <- fin_data %>%
  select(balance, age, day, duration, campaign, previous)

names(regression_df)

# For classification analysis 
classification_df <- fin_data %>%
  mutate(y = factor(y, levels = c("yes", "no"))) %>%
  select(y, age, loan, housing, marital, job, duration, campaign, previous, poutcome, month, balance, age, day, duration, campaign, previous)

# Verify column names and target variable for classification
names(classification_df)
levels(classification_df$y)
```
```{r}
classification_df <- fin_data %>%
  mutate(y = factor(y, levels = c("yes", "no"))) %>%
  select(
    y,
    age,
    loan,
    housing,
    marital,
    job,
    duration,
    campaign,
    previous,
    poutcome,
    month,
    balance,
    day
  )
```

```{r}
summary(fin_data$previous)

sd(fin_data$previous)

n_distinct(fin_data$previous)

table(fin_data$previous)
```
```{r}
anyDuplicated(names(classification_df))

sum(duplicated(names(classification_df)))

sum(duplicated(classification_df))
```


# DATA EXPLORATION
## Row 
### Column {width=100%}
::: {.card title="Data Overview" fill="false"}
* The dataset consists of 45,211 rows and 16 columns, containing a variety of variables as previously described. The dataset was loaded in three different ways: as a tibble for the entire dataset, as a tibble for the regression model, and as a separate tibble for the classification models, as shown below.

:::

### Column {width=40%}
::: {.card title="Fin_data" fill="false"}
```{r}
str(fin_data)
```
:::

### Column {width=30%}
::: {.card title="regression_df" fill="false"}
```{r}
str(regression_df)
```
:::

## Column {width=30%}
::: {.card title="classification_df" fill="false"}
```{r}
str(classification_df)
```
::: 


## Column

::: {.card title="Data Summary" fill="false"}
```{r}
skim(fin_data)
```
:::


## Column {width=50%}

::: {.card title="Dataset Summary" fill="false"}
```{r}
summary(fin_data)
```
:::

## Row
### Column

::: {.card title="Short note on summary" fill="false"}

* We can see that the dataset contains 45,211 observations, with variables including age, job, marital status, education, balance, and other demographic and financial details. The age of the individuals ranges from 18 to 95 years, with a median age of 39, while balance ranges from 0 to 102,127, with a mean of 1,415 and a median of 485. The loan variable is binary (yes/no), and the duration of interactions with clients ranges from 0 to 4,918, with a mean of 258.2 and a median of 180. The poutcome variable and the target y (indicating whether the customer subscribed to a term deposit) are categorical, with the y variable likely showing the outcome of a marketing campaign. 

:::

## Column {width=50%}

::: {.card title="Correlation summary" fill="false"}
```{r}
cor(fin_data[sapply(fin_data, is.numeric)])
```

:::

## Column {width=50%}

::: {.card title="Short note on corelation" fill="false"}

*  The summary of the correlation matrix shows that most of the variables have very low correlations with each other, indicating minimal linear relationships. This suggests that multicollinearity is not a significant concern for regression analysis, and the features are likely independent of one another, which is helpful for building both regression and classification models. 

:::

# DATA VISUALIZATION
## Row


::: {.card title="Histograms" fill="false"}
```{r, fig.align='left'}
#! title: Histograms for Numeric Variables
numeric_df <- fin_data %>% select(where(is.numeric))

numeric_df_long <- numeric_df %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value")


ggplot(numeric_df_long, aes(x = Value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "black") +
  facet_wrap(~Variable, scales = "free") +
  labs(title = "Histograms for Numeric Variables",
       x = "Value",
       y = "Frequency") +
  theme_minimal()
```
:::


::: {.card title="Notes on Histograms" fill="false"}

* **Implications for Regression Models:**

a). Skewed variables may require transformation (e.g., log) to normalize their distribution and prevent model bias.

b). Highly skewed data can cause outliers that affect the accuracy and interpretation of regression coefficients.

c). Non-linearity may arise, requiring adjustments like polynomial features or non-linear models.

* **Implications for Classification Models:**

a). Skewed variables could lead to class imbalance, making the model favor the majority class.

b). Feature transformations or scaling may be necessary to improve model robustness.

c). Resampling techniques or ensemble models can help address imbalance and improve predictions.

:::

## Row

::: {.card title="Correlation Heatmap" fill="false"}
```{r,fig.align='left'}
ggcorrplot(cor(regression_df))
```

:::

::: {.card title="Notes on Correlation Heatmap" fill="false"}

* **The correlation matrix shows strong positive correlations between variables like:**

a). Balance and Age

b). Balance and Duration

c). Previous and Campaign

* **Strong correlations may indicate closely related features, which could lead to multicollinearity issues in regression models.**

* **High multicollinearity can cause:**

a). Instability in regression model coefficients.

b). Reduced interpretability of both regression and classification models.

* **To improve model performance and stability, it may be necessary to:**

a). Remove or combine highly correlated variables.

:::

## Row

::: {.card title="Pairwise Plot" fill="false"}

```{r,fig.align='left'}


numeric_df <- fin_data %>%
  select(age, balance, duration, campaign, previous)

set.seed(42)

GGally::ggpairs(
  sample_n(numeric_df, 2000),
  progress = FALSE
)
```

:::


::: {.card title="Notes on Pairwise Plot" fill="false"}

* **Significant correlations were observed between:**

a). Balance and Duration

b). Day and Campaign

* **Negative correlation between Balance and Campaign (-0.085) and weak positive correlations between other variables suggest that:**

a). Multicollinearity may not be a major issue.

* **However, these relationships should still be considered when building models:**

a). In regression, highly correlated predictors can distort coefficient estimates.

b). In classification, correlated features may affect feature importance and decision boundaries, potentially leading to overfitting if not addressed properly.

:::

## Row

::: {.card title="Bar Plots of categorical variables" fill="false"}

```{r, fig.align='left'}
p7 <- ggplot(aes(x = y), data = classification_df) + 
  geom_bar() + 
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p8 <- ggplot(aes(x = loan), data = classification_df) + 
  geom_bar() + 
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p9 <- ggplot(aes(x = housing), data = classification_df) + 
  geom_bar() + 
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p10 <- ggplot(aes(x = marital), data = classification_df) + 
  geom_bar() + 
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p11 <- ggplot(aes(x = job), data = classification_df) + 
  geom_bar() + 
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

grid.arrange(p7, p8, p9, p10, p11, ncol = 3)



```

:::

```{r, fig.align='right'}
p12 <- ggplot(aes(x = poutcome), data = classification_df) + 
  geom_bar() + 
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p13 <- ggplot(aes(x = month), data = classification_df) + 
  geom_bar() + 
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
grid.arrange(p12, p13, ncol = 3)
```

:::

## Row

::: {.card title="Notes on Bar Plots of Categorical Variables" fill="false"}

* **Main points**

a). y: Target variable imbalance could lead to biased predictions.

b). loan and housing: Imbalanced distributions can also affect predictive accuracy.

c). poutcome: The dominance of the "unknown" category may introduce noise in predictions.

d). job: The large "unknown" category may affect model training and result in less meaningful insights.


* **Implications on Classification Modeling:**

a). Imbalanced data in y, loan, and housing may cause the model to favor the majority class, leading to inaccurate predictions for the minority class.

b). The unknown category in poutcome and job may confuse the model and reduce its effectiveness by introducing noise.

c). The model may overfit to the dominant categories in imbalanced variables, especially in month and poutcome.

d). Classification models may perform poorly on the minority class, especially if class distribution is not considered during model training.

* **Solutions:**

a). Resampling techniques

b). Cross validation

:::




# DATA ANALYSIS

## Row


::: {.card title="Quantitative Analysis" fill="false"}

**Multiple Regression**

Multiple Regression is a statistical technique used to model the relationship between a dependent variable and two or more independent variables. It extends simple linear regression by allowing multiple predictors to influence the target. In order to predict customer balances based on demographic and social factors, we will be performing multiple regression
```{r, text.align="right"}
# Fitting the multiple regression model with specified predictors
reg_spec <- linear_reg() %>% 
  set_engine("lm") %>%
  set_mode("regression") 

# Fitting the model using specific predictors for multiple regression
reg_final_fit1 <- reg_spec %>%
  fit(balance ~ age + campaign + day + duration + previous, data = regression_df)


summary(reg_final_fit1$fit)
car::vif(reg_final_fit1$fit)

```
:::

::: {.card title="Assessing the fit" fill="false"}


```{r}
pred_reg_adv_full1 <- reg_final_fit1$fit %>%
  augment(data = regression_df) %>%
  select(balance, age, campaign, day, duration, .fitted, .resid)

head(pred_reg_adv_full1) %>%
  kable(digits = 3)
```

::: 

## Row

::: {.card title="Model Metrics" fill="false"}

```{r}

pred_reg_adv_full1 %>%
    metrics(truth = balance, estimate = .fitted) %>%
    kable(digits = 3)
```

:::

::: {.card title="Result" fill="false"}

*  Root Mean Squared Error (RMSE) of 3004.647, indicates a relatively high average error in the predictions. The R-squared (rsq) value of 0.010 suggests that the model explains only about 1% of the variance in the data, indicating a poor fit. The Mean Absolute Error (MAE) is 1505.722, which also points to significant average error in predictions.

:::

## Row

::: {.card title="View Residual Diagnostic Plot" fill="false"}

Residual diagnostic plot
```{r}
plot(reg_final_fit1$fit)
```

:::

::: {.card title="Discussion on Residual Diagnostic Plots" fill="false"}

**The diagnostic plots above provide important insights into the model's residuals.**

1. Residuals vs Fitted: The plot suggests non-constant variance (heteroscedasticity) due to the widening of residuals as fitted values increase. This indicates that the model might not be capturing the variance in larger values well, possibly affecting predictions.

2. Q-Q Residuals: The plot shows the standardized residuals deviating from the normal distribution, particularly for extreme values, indicating that the residuals are not normally distributed. This violates one of the assumptions of linear regression, which could impact the model's validity.

3. Scale-Location: This plot also shows heteroscedasticity, with a non-random spread of residuals. It confirms that the variance of residuals is not constant across all levels of fitted values, which can affect the model's accuracy and performance.

4. Residuals vs Leverage: Several points, like the ones at high leverage (around values 42559, 15870, etc.), appear to have high influence on the model. These points are outliers that may disproportionately affect the model's predictions, and further investigation or removal might be necessary.

Overall, the model may not be the best fit therefore adjustments should be made to improve its performance and validity.the model may be struggling with non-constant variance and non-normal residuals, and there are influential outliers that need to be addressed for improving the model.

:::

## Row 

::: {.card title="VIF" fill="false"}


::: {.card title="Comments on VIF" fill="false"}

* The VIF values for all the predictors in the model are close to 1, indicating that there is no significant multicollinearity among the variables. This suggests that each predictor is relatively independent of the others, and the model should not suffer from issues related to multicollinearity.

:::

## Row

::: {.card title="Examining coefficient table" fill="false"}

```{r}
tidy(reg_final_fit1) %>% 
  kable(digits = 3)
```

:::

::: {.card title="Comments on coefficient table" fill="false"}

* all p-values are below 0.05 so we won't have to prune

:::
```{r}
hist(regression_df$balance)
```
```{r}
plot(reg_final_fit1$fit)
```
```{r}
par(mfrow = c(2,2))
plot(reg_final_fit1$fit)
```


```{r}
hist(regression_df$balance)
```


```{r}
plot(reg_final_fit1$fit)
```


```{r}
par(mfrow = c(2,2))
plot(reg_final_fit1$fit)
```


## Row

::: {.card title="Actual vs Predicted plot" fill="false"}
```{r}
pred_reg_final_fit <- reg_final_fit1$fit %>%
                        augment(regression_df) 

#Reminder - if augment with $fit use y=.fitted
#otherwise use .pred (look in object to see which)
pred_reg_final_fit %>%
  ggplot(aes(y = .fitted, x = balance)) + 
    geom_point(col = "#6e0000") +
    geom_abline(col = "gold") + 
    ggtitle("Predicted vs Actual Balance",
            subtitle=paste("Adj R-Squared:",
                          round(glance(reg_final_fit1)$adj.r.squared,2),
                          "  Mean Abs Residuals:",round(mean(abs(reg_final_fit1$fit$residuals)),2))) 
```
:::

::: {.card title="Comment on model" fill="false"}

* The plot shows a poor fit of the model, with the adjusted R-squared value of 0.01 indicating that the model explains only 1% of the variance in the data. The mean absolute residuals of 1505.72 further suggest significant prediction errors, and the data points are scattered widely, especially for higher values of the balance, implying that the model does not capture the relationship well. 

:::

## Row

::: {.card title="Variable Importance Plot" fill="false"}

```{r}
vip(reg_final_fit1)
```
:::

::: {.card title="Comments on vip" fill="false"}

* The plot shows that age is the most important feature in the model, with a significantly higher importance score compared to other variables. The other features such as duration, previous, campaign, and day have much smaller importance, indicating they contribute less to predicting the target variable in this model.

:::

## Row

::: {.card title="Results Metrics" fill="false"}
```{r}
curr_metrics <- reg_final_fit1 %>%
                  augment (regression_df) %>%
                  metrics(truth=balance,estimate=.pred)
results_bank <- tibble(model = 'Multiple Reg',
                  rmse = curr_metrics[[1,3]],
                  rsq=curr_metrics[[2,3]],
                  mae=curr_metrics[[3,3]])
results_bank %>%
  kable(digits = 3)
```

::::

::: {.card title="Comment on metrics" fill="false"}

* The results for the Multiple Regression model show an RMSE of 3004.647 indicating a high average error in predictions. The R-squared (rsq) value is very low at 0.01, suggesting that the model explains almost none of the variance in the data. The MAE of 1505.722 further confirms the model's poor performance in terms of prediction accuracy. This could probably be because of lack of key predictors, or too many variables that are probably adding noise to the model instead signal. However, we are moving forward to another model `Lasso regression` which regularizes the model by penalizing large coefficients, which can help reduce overfitting.

:::

## Row

::: {.card title="Lasso Regression" fill="false"}

* Lasso Regression - Tuned lasso
To improve the model performance and enhance interpretability, we will move forward with lasso regression to know Which customer profile characteristics best predict account `balances` This is useful for our large data set and will reduce overfitting. However before we proceed, we have to go through the following steps below.

* - Regression that prevents overfitting by penalizing large coefficients thus improving model.
* - minimizes the sum of squared errors along with a penalty for the absolute values of the coefficients.
* - simplifies the model by keeping only relevant predictors.
* - improves prediction performance mainly when dealing with many predictors.
* - handles multicollinearity by shrinking correlated variables and setting some of them to zero, making the model more stable.

*  Preprocess Data with `recipe()` & `bake()` -Defining Recipe for Data Preparation and Bake
We will proceed with lasso regression since there is no strong correlation among the variables, although campaign and day tend to have a good correlation but it is not so high that will require eliminating any of these variables. To prevent overfitting, lasso regression will be performed. we are also going to be using tuned lasso to ensure that the model is not too complex, balancing bias and variance, and improving its generalization to new, unseen data.

```{r}
bank_recipe <- recipe(balance ~ ., data = regression_df) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_predictors()) %>%
  prep()
bank_norm <- bake(bank_recipe, new_data = regression_df)
```

Defining a set of metrics
```{r}
my_metrics = metric_set(yardstick::rmse, yardstick::mae, yardstick::rsq)  
```

```{r}
set.seed(42)
bank_split_norm <- initial_split(bank_norm, prop = .60)
bank_train_norm <- rsample::training(bank_split_norm)
bank_test_norm <- rsample::testing(bank_split_norm)

```

```{r}
bank_grid <- tibble(penalty = seq(.01, 100, len = 100))
bank_folds <- vfold_cv(bank_train_norm, v = 5)
```


```{r}
# Define Model Specifications for Lasso Regression
bank_lassotune_spec <- linear_reg(penalty = tune(), 
                                  mixture = 1) %>% 
                       set_engine("glmnet") %>% 
                       set_mode("regression") 

# Create the workflow and tune the penalty, using only quantitative variables
bank_lassotune_wf <- workflow() %>%
                        add_model(bank_lassotune_spec) %>%
                        add_formula(balance ~ age + duration + day + previous + campaign)

# Tune the penalty using resampling
bank_lassotune_rs <- bank_lassotune_wf %>%
                          tune_grid(resamples = bank_folds, 
                                    grid = bank_grid,
                                    metrics = my_metrics)

# Select the best penalty based on RMSE
lowest_rmse_lasso_bank <- bank_lassotune_rs %>%
                       select_best(metric = "rmse")

# Finalize the workflow with the best penalty
final_lasso_bank <- bank_lassotune_wf %>% 
                  finalize_workflow(lowest_rmse_lasso_bank)

# Fit the model on the training data
final_lassobank_fit <- final_lasso_bank %>% 
                       fit(bank_train_norm)

# Make predictions on the test data
pred_final_lassobank_fit <- final_lassobank_fit %>% 
                           augment(bank_test_norm)

# Print the best penalty
print(paste('The lowest rmse Lasso penalty is', lowest_rmse_lasso_bank$penalty))


```
:::

::: {.card title="Lasso regression" fill="false"}

*  Lasso penalty value of 0.01 indicates that the model achieved its best performance in terms of prediction accuracy when the penalty term was set to this value, resulting in the smallest error between the predicted and actual values.

:::

# Row

::: {.card title="Computing Metrics" fill="false"}

```{r}
curr_metrics <- pred_final_lassobank_fit %>%
  my_metrics(truth = balance, estimate = .pred)
results_new <- tibble(model = 'Lasso Tuned',
                  rmse = curr_metrics[[1,3]],
                  mae=curr_metrics[[2,3]],
                  rsq=curr_metrics[[3,3]])
results_bank <- bind_rows(results_bank,results_new)
results_bank %>%
  kable(digits = 3)
```

:::

::: {.card title="Comments" fill="false"}

* The results for the Lasso Tuned model show an RMSE of 3050.961 indicating a high average prediction error, though slightly better than the multiple regression model. The R-squared value of 0.008 is still very low, meaning the model explains almost none of the variance in the data. The MAE of 1505.330 is similar to that of the multiple regression model, suggesting that the model's performance hasn't improved significantly. Thus we will be processing with another called GAMs. 

::: 

## Rows

::: {.card title="Visualizing the values to see the lowest RMSE" fill="false"}

```{r}
df_p <- bank_lassotune_rs %>%
  collect_metrics()
p <- ggplot(df_p, aes(log(penalty), mean, color = .metric))
p + geom_errorbar(aes(ymin = mean - 2*std_err, ymax = mean + 2*std_err),
                  alpha = 0.5) +
  geom_line() +
  facet_wrap(~ .metric, scales = "free", nrow = 2) +
  theme(legend.position = "none")
```
:::


::: {.card title="Comments" fill="false"}

*  The plot above show that as the penalty increases, RMSE and MAE both decrease signifying improved model accuracy, while R-squared also increases suggesting a better fit to the data. However, the plots show that the changes in these metrics plateau as the penalty value grows suggesting that further increasing the penalty might not significantly improve the model.

:::

## Row

::: {.card title="Lasso Regression Model Final Coefficients" fill="false"}


```{r}
final_lassobank_fit %>%
  extract_fit_parsnip() %>% 
  tidy() %>%
  kable(digits = 3)
```
:::

## Row

::: {.card title="VI Plot" fill="false"}
```{r}
final_lassobank_fit %>% 
   extract_fit_parsnip() %>% 
   vip(num_features = 10)
```

:::

::: {.card title="Comments" fill="false"}

* The VI plot illustrates that age is the most significant feature in the model with a significantly higher importance score than the other variables. Previous, duration, campaign, and day have much smaller contributions, suggesting that they have a lesser impact on predicting the target variable compared to age.

## Row

::: {.card title="Coefficients" fill="false"}

```{r}
coef_plot_final <- final_lasso_bank %>% 
  fit(bank_train_norm)

tidy_coefs <- coef_plot_final$fit$fit$fit %>% 
  broom::tidy() %>% 
  filter(term != "(Intercept)") %>% 
  dplyr::select(-step, -dev.ratio)

best_penalty <- tibble(penalty=.01)
delta <- abs(tidy_coefs$lambda - best_penalty$penalty)
lambda_opt <- tidy_coefs$lambda[which.min(delta)]

label_coefs <- tidy_coefs %>% 
  mutate(abs_estimate = abs(estimate)) %>% 
  filter(abs_estimate >= 0.01) %>% 
  distinct(term) %>% 
  inner_join(tidy_coefs, by = "term") %>% 
  filter(lambda == lambda_opt)
```

```{r}
tidy_coefs %>% 
  ggplot(aes(x = lambda, y = estimate, group = term, col = term, label = term)) +
  geom_vline(xintercept = lambda_opt, lty = 3) +
  geom_line(alpha = .4) +
  theme(legend.position = "none") +
  scale_x_log10() +
  ggrepel::geom_text_repel(data = label_coefs, max.overlaps = 100)
```
:::

::: {.card title="Comments" fill="false"}

* The plot illutrates that as lambda increases, the coefficients shrink towards zero, with age remaining the most significant predictor, while the other variables (previous, duration, day, and campaign) have their coefficients reduce signifying less importance in the model as the penalty increases.

## Row

::: {.card title= "Predicted balance vs Actual balance" fill="false"}

```{r}
pred_final_lassobank_fit %>%
  ggplot(aes(y = .pred, x = balance)) + 
      geom_point(col = "#6e0000") +
      geom_abline(col="gold") + 
      ggtitle("Predicted Balance vs Actual Balance",
              subtitle=paste("Lasso Regression","Lambda=",round(lowest_rmse_lasso_bank$penalty,3)))
```

:::

::: {.card title="Comments" fill="false"}

* from the plot above, we can see the spread of points suggests that the model struggles with higher balance values, as the predicted values do not capture the variation at the extremes well, implying potential issues with model fit for those cases. This could be the influence of outliers especially for higher balances, if we remember, the multiple regression analysis shows that the dataset has outliers

## Row

::: {.card title="GAMs" fill="false"}

GAMs
Generalized Additive Models (GAMs) are a flexible type of regression model that allow for non-linear relationships between the independent variables and the dependent variable. Unlike multiple regression, which assumes a linear relationship, GAMs use smooth functions to model complex patterns in the data. This makes GAMs better suited for capturing non-linear trends that simple linear models like multiple regression or Lasso might miss. Compared to Lasso regression, which performs feature selection and shrinks coefficients, GAMs provide more flexibility by allowing each predictor to have its own smooth, non-linear effect, making them more adaptable for complex datasets.
```{r}
# Defining model specification
gam_spec <- gen_additive_mod() %>%
              set_engine("mgcv", importance = "impurity") %>%
              set_mode("regression")
#Fitting the model
gam_fin_fit <- gam_spec %>%
  fit(balance ~ s(age, bs = "cr", k=3) + s(duration, bs = "cr") + s(campaign, bs = "cr", k=3) +
        s(day, bs ="cr", k=5) + s(previous, bs = "cr", k=3),
                data = regression_df)
summary(gam_fin_fit$fit)
```

:::

## Row

::: {.card title="Checking the model diagnostics" fill="false"}

Using `gam.check` to perform some checking of the residuals and k.

```{r}
gam.check(gam_fin_fit$fit)
```
:::

::: {.card title="Comments on GAM Checking" fill="false"}

* Q-Q plot shows that the residuals deviate from a normal distribution, particularly in the tail, which may suggest non-constant variance (heteroscedasticity) or the presence of outliers that could affect the accuracy of the model's predictions. The residuals vs. linear predictor plot indicates a non-linear relationship between the residuals and the fitted values, highlighting potential model misspecification or the need for transformation of variables. The histogram of the residuals further shows that they are skewed, reinforcing concerns about the normality assumption. The response vs. fitted values plot demonstrates a lack of clear linearity, implying that the current model might not capture all the complexity of the data. Overall, these issues suggest that further model refinement, such as addressing outliers or transforming variables, may be necessary for better predictions of balance.

:::


## Row

::: {.card title="RMSE, MAE, AND R-Squared" fill="false"}

Finding error measures (RMSE, MAE, AND R-Squared)
```{r}
pred_gam_fin_fit <- gam_fin_fit %>% 
                    augment(regression_df)
head(pred_gam_fin_fit) %>%
  kable()
```

:::

## Row

::: {.card title="View the RMSE, MAE and R-Squared Results Table" fill="false"}
```{r}
curr_metrics <- pred_gam_fin_fit %>%
                    metrics(truth=balance,estimate=.pred)
results_fin<- tibble(model = 'GAM 1',
                  rmse = curr_metrics[[1,3]], 
                  rsq = curr_metrics[[2,3]],
                  mae = curr_metrics[[3,3]],
                  )
results_fin %>%
  kable(digits = 3)
```

:::

::: {.card title="Comments RMSE, MAE and R-Squared Results Table" fill="false"}

* RMSE of 2998.206, indicates a relatively high average error in predicting the target variable. R-squared value of 0.015 suggests that the model explains very little of the variability in the data, indicating poor predictive power. MAE of 1499.378 further confirms that the model's predictions are far from accurate on average.

:::

## Row 

::: {.card title="View the RMSE, MAE and R-Squared Results Table" fill="false"}

```{r}
ggplot(pred_gam_fin_fit, aes(y=balance, x=.pred)) +
          geom_point(col = "#6e0000") +
          geom_abline(col='orange') +
          ggtitle('GAMS Model')
          

```
:::

::: {.card title="Comment on plot" fill="false"}

* The plot shows a large number of predicted values (.pred) clustered at low values of balance, with a few outliers at higher balance values. The majority of the predictions fall along the lower end of the scale, and the residuals are concentrated near zero, indicating that the model has trouble accurately predicting larger values of balance. This suggests that the model's performance may be limited in capturing higher range outcomes.

:::

## Row

::: {.card title="Plotting GAMs variables vs dependent variable" fill="false"}

```{r}
gam_fin_fit_viz <- getViz(gam_fin_fit$fit)
print(plot(gam_fin_fit_viz, allTerms=TRUE), pages = 1)
```
:::

::: {.card title="Discussion" fill="false"}

* The plot indicates that "age," the relationship with balance appears to be nonlinear and increases exponentially as age increases. The "duration" predictor has a positive linear effect, while "campaign" shows a slight upward trend, with some higher variation for large values. "Day" shows a cyclical pattern, with significant peaks around certain days. Lastly, "previous" shows a steep drop after a certain threshold, suggesting a diminishing effect as the number of previous contacts increases. 

:::

## Row

::: {.card title="Discussion" fill="false"}

```{r}
curr_metrics <- pred_gam_fin_fit %>%
  my_metrics(truth = balance, estimate = .pred)
results_new <- tibble(model = 'GAMs',
                  rmse = curr_metrics[[1,3]],
                  mae=curr_metrics[[2,3]],
                  rsq=curr_metrics[[3,3]])
results_bank <- bind_rows(results_bank,results_new)
results_bank %>%
  kable(digits = 3)
```

:::

::: {.card title="Discussion" fill="false"}

* The results shows that GAMs has the best performance among the three, with the lowest RMSE of 2998.206 and a slightly better R-squared value of 0.015 compared to the other two models. This suggests that the GAM model is able to explain more of the variance in the target variable, "balance," than both the multiple regression and lasso tuned models. The multiple regression model has an RMSE of 3004.647, which is very close to the GAM model, but its R-squared value is much lower (0.010), indicating it doesn't fit the data as well. The lasso-tuned model has an RMSE of 3050.961, which is the highest among the three models, and its R-squared is the lowest at 0.008, suggesting it has the poorest fit to the data. However, the MAE for all models is quite close to each other, with GAMs yielding the smallest value (1499.378), followed by multiple regression (1505.722) and lasso-tuned (1505.330). Overall, while all models have similar MAE values, the GAM outperforms both the multiple regression and lasso-tuned models in terms of RMSE and R-squared, making it the most effective model for predicting the balance in this dataset.

:::

## Row

** Quantitative Analysis

::: {.card title="Random Forest Model" fill="false"}

Random Forest Model
Lets generate a table for y to know the counts for each binary value
```{r}
table(classification_df$y)
```
:::

::: {.card title="Counts" fill="false"}

* We can see we have a lot of no(s) compared to the number of yes(s) which means the data is not balanced and we cannot use it as our target variable like this so we have to preprocesses the dataset by normalizing, encoding, removing zero-variance predictors, and balancing the class distribution before returning the final transformed dataset. We used the `bake()`, `recipe()` and `smote()` functions for this purpose. 

:::

## Row

::: {.card title="Random Forest Model" fill="false"}

```{r}
fin_recipe <- recipe(y ~ ., data = classification_df) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_smote(y, over_ratio = .5) %>%
  prep()
df_fin_baked <- bake(fin_recipe, new_data = NULL)

```

:::

## Row

::: {.card title="Counts for y variable" fill="false"}

```{r}
table(df_fin_baked$y)
```
:::

## Row

::: {.card title="Data Training" fill="false"}

Splitting the data into a 60% training set and a test set.
```{r}
set.seed(2934)
df_fin_baked_split <- rsample::initial_split(df_fin_baked, prop=.6, strata=y)
df_fin_baked_train <- rsample::training(df_fin_baked_split)
df_fin_baked_test <- rsample::testing(df_fin_baked_split)

```

:::

## Row

::: {.card title="Fitting Tuned Random Forest Model" fill="false"}

Fitting a Tuned Random Forest model
```{r}
# Defining the CV folds and grid parameters
set.seed(234)
fin_folds <- vfold_cv(df_fin_baked_train, v = 3)
finrf_grid <- expand_grid(mtry = seq(5, 15, by = 5),
                       trees = c(100, 200))
#Defining the Model specifications
finrf_tune_spec <- rand_forest(mtry = tune(), trees = tune()) %>%
  set_engine("ranger", importance = "impurity") %>% 
  set_mode("classification") 
#Defining our workflow 
finrf_wf <- workflow() %>%
  add_model(finrf_tune_spec) %>%
  add_formula(y ~ .)

#Tuning on the grid of values
finrf_rs <- finrf_wf %>% 
  tune_grid(resamples = fin_folds,
            grid = finrf_grid)

```

:::

## Row

::: {.card title="Finalize with the best metrics" fill="false"}

```{r}
#Finalizing the workflow with the best model
final_finrf_wf <- 
  finrf_wf %>% 
  finalize_workflow(select_best(finrf_rs))
final_finrf_wf

```
:::

## Row

Model parameters in final model
```{r}
  mtry = 19
 trees = 1000
#Fit the final model to the training data
final_finrf_tune <- 
  final_finrf_wf %>%
  fit(data = df_fin_baked_train) 
final_finrf_tune

```
:::

## Row

Confusion matrix
```{r}
pred_finrf_tune <- final_finrf_tune %>%
          augment(df_fin_baked_test) 
pred_finrf_tune %>%
  yardstick::conf_mat(truth = y, estimate = .pred_class)
  
```
::: 

## Row

::: {.card title="Random Forest Model - Metrics" fill="false"}
```{r}
metrics <- metric_set(yardstick::accuracy, yardstick::sensitivity, yardstick::specificity, yardstick::precision)

curr_metrics <- pred_finrf_tune %>%
  metrics(truth=y,estimate=.pred_class)
results_fin <- tibble(model = 'Tuned RF Model',
                      accuracy = curr_metrics[[1,3]], 
                      sensitivity = curr_metrics[[2,3]],
                      specificity = curr_metrics[[3,3]],
                      precision = curr_metrics[[4,3]],
                      avg_sens_prec = (sensitivity+precision)/2
                      )
results_fin %>%
  kable()

```

:::

::: {.card title="Random Forest Model Metrics" fill="false"}

* As we can see, the model performs well with an accuracy of 91.8%. It achieves a sensitivity of 86.5%, meaning it correctly identifies 86.5% of positive cases, and a specificity of 94.4%, meaning it correctly identifies 94.4% of negative cases. The precision of 88.6% indicates that when the model predicts a positive case, it is correct 88.6% of the time, and the average of sensitivity and precision is 87.6%.

:::

## Row

::: {.card title="Random Forest Model VIP" fill="false"}

```{r}
final_finrf_tune %>% 
  extract_fit_parsnip() %>% 
  vip()
```
:::

::: {.card title="Notes on Random Forest Model" fill="false"}

* From the plot, we can that the most important feature in predicting the target variable is `duration`, followed by `campaign` and `poutcome_success`, which indicates that the length of the interaction and previous campaign activity are critical for the subscription to the product.Although, other important features include `housing_yes`," `age`, and `day` while `month_may` and `poutcome_unknown` have relatively less influence on individual's subscription to the product. This shows that subscribing to a product is highly influenced by recent campaign activity and duration of engagement.

:::

## Row

::: {.card title="Random Forest Model - ROC Curve" fill="false"}
```{r}
#DO NOT EDIT FUNCTION
ROC_graph <- function(pred_data, truth, probs, model_desc="", df_roc = ""){
    #This function creates a ROC Curve. It will return a df_roc with values
    #it used to create the graph. It will also add to a previous ROC curve
    #The inputs are the prediction table (from augment()) and the columns for the
    #truth and probability values. There is also an optional model description
    #and a previous df_roc dataframe. The columns need to be strings (i.e., 'sales')
    #Capture the auc value
    curr_auc <- pred_data %>%
                     roc_auc(truth = {{truth}}, {{probs}}) %>%
                     pull(.estimate)
    #Capture the thresholds and sens/spec
    ###First choice creates a new df_roc table
    if (mode(df_roc) == "character") { #if it is a tibble will be "list"
        df_roc <- pred_data %>% roc_curve(truth = {{truth}}, {{probs}}) %>% 
                          mutate(model = paste(model_desc,round(curr_auc,2)))
    }
    ###Second choice is if there is already a df_roc that was input
    else {
    df_roc <- bind_rows(df_roc,  #use if df_roc exists with other models
                   pred_data %>% roc_curve(truth = {{truth}}, {{probs}}) %>% 
                      mutate(model = paste(model_desc,round(curr_auc,2))))
    }
    #Plot the ROC Curve(s) 
    print(ggplot(df_roc, 
            aes(x = 1 - specificity, y = sensitivity, 
                group = model, col = model)) +
            geom_path() +
            geom_abline(lty = 3)  +
            scale_color_brewer(palette = "Dark2") +
            theme(legend.position = "top") )
    #Capture the roc values in a df to add additional ROC curves
    return(df_roc)
}
    
df_roc_fin <- ROC_graph(pred_finrf_tune,"y",".pred_yes","Tuned RF")

```
:::

::: {.card title="Notes Random Forest Model ROC Curve" fill="false"}

* As we can see, from the ROC curve, the model has a high sensitivity (true positive rate) and low false positive rate (1-specificity) indicating it performs very well in identifying positive instances. The curve is steep, indicating excellent classification ability, and the model has a high area under the curve (AUC) value of 0.98, which reflects its strong discriminatory power between the two classes.

:::

## Row

::: {.card title="Random Forest Model- Threshold Bestcutoff" fill="false"}

```{r}
#Find Best Threshold cutoff (DO NOT EDIT FUNCTION)
ROC_threshold <- function(pred_data,truth,probs) {
  #This function finds the cutoff with the max sum of sensitivity and specificity
  #Created tidy version of:
  #http://scipp.ucsc.edu/~pablo/pulsarness/Step_02_ROC_and_Table_function.html
  #The inputs are the prediction table (from augment()) and the columns for the
  #truth and probability values. The columns need to be strings (i.e., 'sales')
 
  roc_curve_tbl <- pred_data %>% 
                    roc_curve(truth = {{truth}}, {{probs}}) 
  auc = pred_data %>%
              roc_auc(truth = {{truth}}, {{probs}}) %>%
              pull(.estimate)
  best_row = which.max(roc_curve_tbl$specificity + roc_curve_tbl$sensitivity)
  return(tibble(Best_Cutoff = round(pull(roc_curve_tbl[best_row,'.threshold']),4),
         Sensitivity = round(pull(roc_curve_tbl[best_row,'sensitivity']),4),
         Specificity =round(pull(roc_curve_tbl[best_row,'specificity']),4),
         AUC_for_Model = round(auc,4)))
}
set.seed(123)
best_cut_finrf <- ROC_threshold(pred_finrf_tune, 'y', '.pred_yes')
best_cut_finrf %>%
  kable(digits=3)

```

:::

::: {.card title="Random Forest Model Threshold Bestcutoff" fill="false"}

* The best cutoff for the model is `0.337` and it achieves a high sensitivity of `94.7%`, showing it correctly identifies most positive cases. The specificity of 88.9% indicates it also correctly identifies negative cases and with an AUC of 0.976, the model demonstrates excellent overall performance in distinguishing between the two classes (yes(s) and no(s).

:::

```{r}
#Create the new column
pred_finrf_tune <- pred_finrf_tune %>%
                   mutate(pred_y_newcut = factor(ifelse(.pred_yes > best_cut_finrf[[1]],"yes","no"),
                                             levels=c("yes","no")))
#View the confusion matrix
pred_finrf_tune %>% 
  yardstick::conf_mat(truth = y, estimate = pred_y_newcut)

```
:::

## Row

::: {.card title="Random Forest Model Metrics on Bestcutoff" fill="false"}
```{r}
metrics <- pred_finrf_tune %>%
  metrics(truth=y,estimate=pred_y_newcut)
results_new<- tibble(model = paste('Tuned RF Model Cutoff',round(best_cut_finrf[[1]],2)),
                      accuracy = curr_metrics[[1,3]], 
                      sensitivity = curr_metrics[[2,3]],
                      specificity = curr_metrics[[3,3]],
                      precision = curr_metrics[[4,3]],
                      avg_sens_prec = (sensitivity+precision)/2)
results_fin <- bind_rows(results_fin, results_new)
results_fin %>%
  kable()

```

::: 

## Row

::: {.card title="Fitting a tuned Gradient Boosted Tree Model" fill="false"}



```{r}
#Define the CV folds and grid parameters
set.seed(3345)
fin_boost_grid <- grid_space_filling(
  tree_depth(),
  loss_reduction(),
  learn_rate(),
  finalize(mtry(), df_fin_baked_train),
  size = 10)

#Define the Model specifications
fin_boost_tune_spec <- boost_tree(trees = 200,
                              min_n = 5,
                              tree_depth = tune(),
                              learn_rate = tune(),
                              mtry = tune(),
                              loss_reduction = tune()) %>%
  set_engine("xgboost") %>% 
  set_mode("classification")

#Define our workflow 
fin_boost_wf <- workflow() %>%
  add_model(fin_boost_tune_spec) %>%
  add_formula(y ~ .)

#Tune on the grid of values
fin_boost_rs <- fin_boost_wf %>% 
  tune_grid(resamples = fin_folds,
            grid = fin_boost_grid,
            control = control_grid(save_pred = TRUE))

```

Finalize with the best metrics
```{r}
#Finalize the workflow with the best model
final_fin_boost_wf <- 
  fin_boost_wf %>% 
  finalize_workflow(select_best(fin_boost_rs))
final_fin_boost_wf


```
Model Parameters in final model
```{r}
mtry = 22
  trees = 200
  min_n = 5
  tree_depth = 11
  learn_rate = 4.3329540878864e-07
  loss_reduction = 0.000164175174703485
  
#Fit the final model to the training data
final_boost_tune <- 
  final_fin_boost_wf %>%
  fit(data = df_fin_baked_train) 
final_boost_tune

```
Confusion Matrix
```{r}
pred_finboost_tune <- final_boost_tune %>%
          augment(df_fin_baked_test) 
pred_finboost_tune %>%
  yardstick::conf_mat(truth = y, estimate = .pred_class)

```
Result Table with added metrics
```{r}
metrics <- pred_finboost_tune %>%
  metrics(truth=y,estimate=.pred_class)
results_new<- tibble(model = 'Tuned Boost Model',
                      accuracy = curr_metrics[[1,3]], 
                      sensitivity = curr_metrics[[2,3]],
                      specificity = curr_metrics[[3,3]],
                      precision = curr_metrics[[4,3]],
                      avg_sens_prec = (sensitivity+precision)/2)
results_fin <- bind_rows(results_fin, results_new)
results_fin %>%
  kable()

```
:::

## Row

::: {.card title="Fitting a tuned Gradient Boosted Tree Model - ROC Curve" fill="false"}

```{r}
df_roc_fin <- ROC_graph(pred_finboost_tune,"y",".pred_yes","Boosted", df_roc_fin)
```
:::

## Row

::: {.card title=" tuned Gradient Boosted Tree Model- VIP" fill="false"}

```{r}
final_boost_tune  %>% 
  extract_fit_parsnip() %>%
  vip()

```
::: 

## Row

Best cut off
```{r}
set.seed(123)
best_cut_boost <- ROC_threshold(pred_finboost_tune, 'y', '.pred_yes')
best_cut_boost %>%
  kable(digits=3)

```
Creating a new column
```{r}
#Create the new column
pred_finboost_tune <- pred_finboost_tune %>%
                   mutate(pred_y_newcut2 = factor(ifelse(.pred_yes > best_cut_boost[[1]],"yes","no"),
                                             levels=c("yes","no")))
#View the confusion matrix
pred_finboost_tune %>% 
  yardstick::conf_mat(truth = y, estimate = pred_y_newcut2)

          
```
:::

## Row

::: {.card title="Adding metrics to the results table" fill="false"}

```{r}
metrics <- pred_finboost_tune %>%
  metrics(truth=y,estimate=pred_y_newcut2)
results_new<- tibble(model = paste('Tuned Boosted Model Cutoff',round(best_cut_boost[[1]],2)),
                      accuracy = curr_metrics[[1,3]], 
                      sensitivity = curr_metrics[[2,3]],
                      specificity = curr_metrics[[3,3]],
                      precision = curr_metrics[[4,3]],
                      avg_sens_prec = (sensitivity+precision)/2)
results_fin <- bind_rows(results_fin, results_new)
results_fin %>%
  kable(digits = 3)

```

:::

## Row

::: {.card title="Logistic Regression" fill="false"}
```{r}
log_reg_spec <- logistic_reg() %>% 
                  set_engine("glm") %>% 
                  set_mode("classification")
log_reg_fit <- log_reg_spec %>%
                  fit(y ~ ., data = df_fin_baked_train)
```

Viewing Confusion matrix
```{r}
pred_log_reg <- log_reg_fit %>% 
                  augment(df_fin_baked_test) 
pred_log_reg %>%
          conf_mat(truth=y,estimate=.pred_class)

```
Viewing Metrics
```{r}
metrics <- pred_log_reg %>%
 metrics(truth = y, estimate = .pred_class)

results_new<- tibble(model = 'Logistic Model',
                     accuracy = curr_metrics[[1,3]], 
                     sensitivity = curr_metrics[[2,3]],
                     specificity = curr_metrics[[3,3]],
                     precision = curr_metrics[[4,3]])
results_fin <- bind_rows(results_fin,results_new)
results_fin %>%
  kable(digits =2)
```

:::

::: {.card title="Result Metrics" fill="false"}

## Row

::: {.card title="ROC Curve for the 3 models" fill="false"}
```{r}
df_roc_fin <- ROC_graph(pred_log_reg,"y",".pred_yes","logistic", df_roc_fin)
```
:::



# DISCUSSION AND CONCLUSION
## Row


::: {.card title="Discussion" fill="false"}

**Discussion and Conclusion**

Based on the results for the regression models, GAMs seems to be the best model. It has the lowest RMSE (2998.206), which indicates it makes more accurate predictions compared to the Multiple Regression and Lasso Tuned models. Although the R-squared (rsq) values for all models are quite low, GAMs still slightly outperforms the others, suggesting that it captures more variance in the data. Additionally, MAE for GAMs (1499.378) is slightly lower than the others, further supporting its better predictive performance. Therefore, GAMs is the best option for this data set even though it didnt capture the data set appropriately.

Moving forward to the classification models, the ROC curve and the performance metrics for three models (Tuned Random Forest, Tuned Boosted, and Logistic Regression) show similar performance, with an accuracy of 0.92, sensitivity and specificity around 0.87 and 0.94, respectively, and precision of 0.89. The ROC curve indicates that both the Tuned Random Forest and Tuned Boosted Models achieve almost identical performance, with a near-perfect curve, suggesting strong ability to distinguish between subscribers and non-subscribers. The Logistic Regression model, while also good, has a slightly lower curve compared to the other two, indicating it might not perform as well in identifying patterns in the data. Given the performance metrics and the ROC curve, the Tuned Random Forest and Tuned Boosted Models are likely the best at predicting what makes an individual subscribe to the product, with their slightly higher overall performance.

Overall, age appears to be the most significant predictor of customers' account balances. However, the actual vs. predicted plot shows a large number of predicted values clustered at low balance values, with a few outliers at higher balances. The majority of the predictions fall at the lower end of the scale, and the residuals are concentrated near zero, suggesting that the model struggles to accurately predict larger balance values. This indicates that the model's performance may be limited in capturing higher-range outcomes.
Based on the variable importance (VI) plots from both Lasso Regression and Multiple Regression, as well as the coefficient plots, age consistently emerges as a key predictor, despite the very poor adjusted R-squared values for the regression models. These poor results may be due to the limited set of predictors used, and the model's performance might improve with the inclusion of other key predictors or through data transformation.
Regarding the classification models, duration and campaign seem to be the most significant predictors of whether an individual will subscribe to a product, while month appears to be the least significant predictor in determining subscription likelihood. 